---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---

2.	Use German Credit Scoring data. (NOT the data we used for lecture notes.)

The German credit score data are downloadable from http://archive.ics.uci.edu/ml/datasets/Statlog+(German+Credit+Data)

(i)	Random sample a training data set that contains 70% of original data points. Start with exploratory data analysis. Find a best model for Credit Scoring data using logistic regression with AIC and BIC. Draw ROC curve, report the AUC, and present the misclassification rate table of your final model. 

(ii)	Test the out-of-sample performance. Using final logistic linear model built from (i) on the 70% of original data, test with the remaining 30% testing data.  (Try predict() function in R.) Report out-of-sample AUC and asymmetric misclassification rate. 
Write a brief report up to 8 pages including all plots and tables and interpret your results.  

```{r}
library(dplyr)
library(ggplot2)
library(ROCR)
```

```{r}
german_credit = read.table("http://archive.ics.uci.edu/ml/machine-learning-databases/statlog/german/german.data")

colnames(german_credit)=c("chk_acct",
                        "duration",
                        "credit_his",
                        "purpose",
                        "amount",
                        "saving_acct",
                        "present_emp",
                        "installment_rate",
                        "sex",
                        "other_debtor",
                        "present_resid",
                        "property",
                        "age",
                        "other_install",
                        "housing",
                        "n_credits",
                        "job",
                        "n_people",
                        "telephone",
                        "foreign",
                        "response")


#orginal response coding 1= good, 2 = bad
#we need 0 = good, 1 = bad
german_credit$response = german_credit$response - 1
```

```{r}
glimpse(german_credit)
```


**Basic EDA**

**Summary statistics**
```{r}
summary(german_credit)
```

**Univariate Analysis**
```{r}
#Function to plot histogram
plot_bar_dodge <- function(x)
{
  ggplot(german_credit, aes(x=x, fill=as.factor(response))) +
  geom_bar( position = 'dodge') +
  xlab(names(x))
}
```

```{r}
plot_hist_dodge <- function(x)
{
  ggplot(german_credit, aes(x=x, fill=as.factor(response))) +
  geom_histogram( position = 'dodge') +
  xlab(names(x))
}

```

```{r}
plot_hist_density <- function(x)
{
  ggplot(german_credit, aes(x = x, fill = as.factor(response))) + 
    geom_density(alpha=.3)
}
```

```{r}
##Function to plot box plot
plot_box <- function(x)
{
  ggplot(german_credit, aes(x = as.factor(response), y = x, fill = as.factor(response))) +
    geom_boxplot()
}
```

**Categorical Variables**
```{r}
cat_var <- c("chk_acct",
             "credit_his",
             "purpose",
             "saving_acct",
             "present_emp",
             "sex",
             "other_debtor",
             "property",
             "other_install",
             "housing",
             "job",
             "telephone",
             "foreign")
```


Proportion table/Contingency table -- To fix
```{r}
prop_split <- function(x)
{
  with(german_credit, table(response, x)) %>% prop.table()
}
```

```{r}
for(i in cat_var)
{
  print(prop_split(german_credit[,i]))
}
```


Histograms for categorical variables
```{r}
for(i in cat_var)
{
  print(plot_bar_dodge(german_credit[,i]))
}
```


**Numerical variables**

1. duration
2. amount
3. installment.rate
4. present_resid
5. age
6. n_credits
7. n_people

Visualizing the distribution of these variables against the response variable


**1. Duration**
```{r}
plot_hist_density(german_credit$duration)
plot_box(german_credit$duration)
```

**2. Amount**
```{r}
plot_box(german_credit$amount)
```

**3. Installment Rate**
```{r}
plot_bar_dodge(german_credit$installment_rate)
```

**4. Residence Since**
```{r}
plot_bar_dodge(german_credit$present_resid)
plot_box(german_credit$present_resid)

```

**5. Age**
```{r}
plot_box(german_credit$age)
```

**6. Credits**
```{r}
plot_box(german_credit$n_credits)
```

**7. n_people**
```{r}
plot_box(german_credit$n_people)
plot_hist_dodge(german_credit$n_people)
```



```{r}
set.seed(13437586)
```

**Train/Test split**
Randomly split the data to training (70%) and testing (30%) datasets:

```{r}
index <- sample(nrow(german_credit),nrow(german_credit)*0.70)
credit.train = german_credit[index,]
credit.test = german_credit[-index,]
```

```{r}
glimpse(credit.train)
```

**Find a best model for Credit Scoring data using logistic regression with AIC and BIC.**

Training the model with all the predictor variables
```{r}
credit.glm0 <- glm(response~., family=binomial, data=credit.train)
```

Extracting certain criteria from the zero-model
```{r}
cat("Deviance = ",round(credit.glm0$deviance,2))
cat("\nAIC = ",round(AIC(credit.glm0),2))
cat("\nBIC = ",round(BIC(credit.glm0),2))
```

Variable selection with stepwise approach

AIC based model
```{r}
credit.aic <- step(credit.glm0, trace = 0)
summary(credit.aic)
```

glm(formula = response ~ chk_acct + duration + credit_his + purpose + 
    amount + saving_acct + installment_rate + sex + other_debtor + 
    other_install + housing + foreign, family = binomial, data = credit.train)

Significant predictors -
1. chk_acct
2. duration
3. credit_his
4. purpose
5. amount
6. saving_acct
7. installment_rate
8. sex
9. other_debtor
10. other_install
11. housing
12. foreign

```{r}
cat("Deviance = ",round(credit.aic$deviance,2))
cat("\nAIC = ",round(AIC(credit.aic),2))
cat("\nBIC = ",round(BIC(credit.aic),2))

```

BIC based model
```{r}
credit.bic <- step(credit.glm0, trace = 0, k=log(nrow(credit.train)))
summary(credit.bic)
```

glm(formula = response ~ chk_acct + duration + other_debtor + 
    foreign, family = binomial, data = credit.train)

Significant predictors are -
1. chk_acct
2. duration
3. other_debtor
3. foreign


```{r}
cat("Deviance = ",round(credit.bic$deviance,2))
cat("\nAIC = ",round(AIC(credit.bic),2))
cat("\nBIC = ",round(BIC(credit.bic),2))
```

Given the AIC of the model based on AIC stepwise approach is smaller. We will select the AIC based model

Final model
```{r}
credit.model <- glm(formula = response ~ chk_acct + duration + credit_his + purpose + 
    amount + saving_acct + installment_rate + sex + other_debtor + 
    other_install + housing + foreign, family = binomial, data = credit.train)


summary(credit.model)
```

```{r}
cat("Deviance = ",round(credit.model$deviance,2))
cat("\nAIC = ",round(AIC(credit.model),2))
cat("\nBIC = ",round(BIC(credit.model),2))
```


**Draw ROC curve, report the AUC, and present the misclassification rate table of your final model.**

```{r}
pred.model.train<- predict(credit.model, type="response")
```


```{r}
par(pty = "s")
pROC::roc(credit.train$response, credit.model$fitted.values, legacy.axes=TRUE, 
          xlab="False Positive Rate", ylab="True Postive Rate",plot=TRUE, col="#377eb8", lwd=4, print.auc = TRUE)

```


**Misclassification rate table**

```{r}
pred.model.train<- predict(credit.model, type="response")
```

**Naive Choice of Cut-off probability**
```{r}
pcut1<- mean(credit.train$response)

# get binary prediction
class.model.train<- (pred.model.train > pcut1)*1

# get confusion matrix
table(credit.train$response, class.model.train, dnn = c("True", "Predicted"))
```


**Determine Optimal cut-off Probability using Cost Function**
```{r}
costfunc = function(obs, pred.p, pcut){
    weight1 = 5   # define the weight for "true=1 but pred=0" (FN)
    weight0 = 1    # define the weight for "true=0 but pred=1" (FP)
    c1 = (obs == 1)&(pred.p < pcut)    # count for "true=1 but pred=0"   (FN)
    c0 = (obs == 0)&(pred.p >= pcut)   # count for "true=0 but pred=1"   (FP)
    cost = mean(weight1*c1 + weight0*c0)  # misclassification with weight
    return(cost) # you have to return to a value when you write R functions
} # end of the function
```


```{r}
p.seq = seq(0.01, 1, 0.01) 
```


```{r}
cost = rep(0, length(p.seq))  
for(i in 1:length(p.seq)){ 
    cost[i] = costfunc(obs = credit.train$response, pred.p = pred.model.train, pcut = p.seq[i])  
} # end of the loop
```

```{r}
plot(p.seq, cost)
```

```{r}
optimal.pcut = p.seq[which(cost==min(cost))]
optimal.pcut
```

```{r}
class.train.opt <- (pred.model.train > optimal.pcut)*1

# step 2. get confusion matrix, MR, FPR, FNR
table(credit.train$response, class.train.opt, dnn = c("True", "Predicted"))
```
```{r}
MR<- mean(credit.train$response!= class.train.opt)
FPR<- sum(credit.train$response==0 & class.train.opt==1)/sum(credit.train$response==0)
FNR<- sum(credit.train$response==1 & class.train.opt==0)/sum(credit.train$response==1)
cost<- costfunc(obs = credit.train$response, pred.p = pred.model.train, pcut = optimal.pcut) 
```

```{r}
cat("Misclassification Rate = ", MR)
cat("\nFalse Positive Rate = ", FPR)
cat("\nFalse Negative Rate = ", FNR)
cat("\nCost = ", cost)
```

**(ii)	Test the out-of-sample performance. Using final logistic linear model built from (i) on the 70% of original data, test with the remaining 30% testing data.**

**Out of sample Prediction**
```{r}
pred.model.test <- predict(credit.model, newdata = credit.test, type = "response")
```

**ROC Curve and AUC**
```{r}
par(pty = "s")
pROC::roc(credit.test$response, pred.model.test, legacy.axes=TRUE, 
          xlab="False Positive Rate", ylab="True Postive Rate",plot=TRUE, col="#377eb8", lwd=4, print.auc = TRUE)
```


```{r}
class.test <- (pred.model.test > optimal.pcut)*1

# step 2. get confusion matrix, MR, FPR, FNR
table(credit.test$response, class.test, dnn = c("True", "Predicted"))
```

```{r}
MR <- mean(credit.test$response!= class.test)
FPR <- sum(credit.test$response==0 & class.test==1)/sum(credit.test$response==0)
FNR <- sum(credit.test$response==1 & class.test==0)/sum(credit.test$response==1)
cost <- costfunc(obs = credit.test$response, pred.p = pred.model.test, pcut = optimal.pcut) 
```

```{r}
cat("Misclassification Rate = ", MR)
cat("\nFalse Positive Rate = ", FPR)
cat("\nFalse Negative Rate = ", FNR)
cat("\nCost = ", cost)
```

